---
title: "00_sp_complier"
author: "Caroline McKeon"
date: "03/17/2020"
output: word_document
---

This markdown is for the extraction, wrangling and analysis of data relating to species spatial pattern (range metrics) and climate and humanfootprint.

# Set up
```{r setup, include=FALSE}
#install.packages(c("tidyverse", "glmmTMB"))
#install.packages("RPostgreSQL", type="source") 
#install.packages("BIEN") 
source("00_sp_functions.R")
knitr::opts_chunk$set(include=FALSE) 
knitr::opts_chunk$set(fig.width=10, fig.height=7) 

## Quick look at model dataframe
## Factors
# for (i in names(Filter(is.factor, ModelDF))) {
#   plot(ModelDF[,i], 
#        main = paste(i))
# }
# 
# ## Numeric variables
# for (i in names(Filter(is.numeric, ModelDF))) {
#   hist(ModelDF[,i], 
#        breaks = 3000, 
#        main = paste(i), 
#        xlab = paste(i))
# }
```

## get range metrics
Read in and clean species range metric data
```{r}
metrics <- read.csv("GRSmetrics_AFE_pol_50km_pol_line_mcp.csv") ## metrics provided by Anna Csergo in Spring 2022. 4260 obs of 42 vars
metrics <- metrics[metrics$Model == "Occurrence",] ## 810 species
metrics <- metrics[which(!is.na(metrics$Species)),]  ## 809 species
metrics$Species <- gsub(" ssp.*", "", metrics$Species) ## remove subspecies (for now). 714 unique species
metrics$Species <- gsub("[[:space:]]*$", "", metrics$Species) ##  692 unique species
small <- c(" s.lat.", " s.str.")
metrics$Species <- gsub(paste(small, collapse="|"), "", metrics$Species) ## 691
metrics$Species <- gsub("//.", "", metrics$Species) 

# Quercus petraea.
# Ranunculus montanus.
# Pinus uncinata.

metrics <- unique(metrics) # 805 obs of 42 vars
metrics <- unique(metrics[, which(names(metrics) %in% c("Species", "total.area", "range.size", "effective.mesh.size", 
                                                        "mean.shape.index", "prop.landscape", "perimeter.area.frac.dim"))])
metrics <- metrics[c("Species", "total.area", "range.size", "effective.mesh.size", "prop.landscape","mean.shape.index", "perimeter.area.frac.dim")]
metrics <- unique(metrics) # 810 obs of 7 vars
names(metrics) <- c("species","Occupied area", "Geographic range size", "Patch size distribution", "Geographic range filling", "Patch shape complexity", "Geographic range fractality") 

## tidy names
metrics$species[grep("oslo", metrics$species, useBytes = TRUE)] <- "Saxifraga osloensis"
metrics$species[metrics$species == "Salix repens coll."] <- "Salix repens"

metrics$species <- factor(metrics$species)
levels(metrics$species) <- gsub(" ", "_", levels(metrics$species))
```

```{r}
## Numeric variables
for (i in names(Filter(is.numeric, metrics))) {
  hist(metrics[,i],
       breaks = 100,
       main = paste(i),
       xlab = paste(i))
}
```
## get phylogeny
Set up phylogeny
```{r}
if(!exists("clean_tree")) {
  try(clean_tree <- read.tree("Data_sp_clean_tree.tre"))
  } else source("01_clean_phylogeny.R")

## 653 species with metrics and phylogeny April 2022
```


## get current climate, humanfootprint and climate change velocity means across all species total areas
```{r}
## load base rasters and occurrence data
#source("02_remote_a_load_hf_vel_clim.R")

## for all species make 1km2 resolution occurrence rasters based on the AFE irregular grids cells 
#source("02_remote_b_make_rasters.R")
#source("02_remote_c_make_rasters.R")

## get mean values for hf, vel and climate across each species' occurrence
#source("02_remote_d_get_ratios.R")

```

have a quick look at everything

```{r}
rat <- readRDS("Data_ratios_dataframe.rds")

for(i in names(Filter(is.numeric, rat))) {
  hist(rat[,i],
       breaks = 50,
       main = paste(i),
       xlab = paste(i))
}
## is these look different from previous histograms that is fine. I have check the values against the rasters and i trust that the 1km2 naming is correct.

```

run analyses
```{r}
#source("03_sp_mean_analysis.R")
#source("03_sp_range_analysis.R")
```


```{r}
# current vs historic calc.
mat_var <- raster("wc2/wc2.1_30s_bio_4.tif") ## mean annual temp SD*100
current_historic <- brick(vel,mat_var)
#current_historic <- drop_na(data.frame(raster::extract(current_historic,sp_co)))
current_historic <- drop_na(data.frame(current_historic))
cor(scale(current_historic$Velocity), scale(current_historic$wc2.1_30s_bio_4)) ## correlation = 0.382153
plot(scale(current_historic$wc2.1_30s_bio_4), scale(current_historic$Velocity))

x <- lm(scale(log(current_historic$Velocity)) ~ scale(log(current_historic$wc2.1_30s_bio_4)))
summary(x)

par(mfrow = c(1,2))
hist(log(current_historic$Velocity))
hist(log(current_historic$wc2.1_30s_bio_4))


grid <- shapefile("AFEcells/cgrs_grid.shp")


```






