---
title: "04_sp_mean_results"
author: "Caroline McKeon"
date: "5/18/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height = 10, fig.width = 16)
source("00_sp_functions.R")
```

```{r plotting themes}
# blank theme
set_theme(
  base = theme_classic(),
  axis.title.size = 2,
  axis.textsize = 1.2,
  legend.size = 2,
  legend.title.size = 2,
  geom.label.size = 3,
  plot.margin = margin(3, 5, 1, 5, "pt")
)

## text editing
get_wraper <- function(width) {
  function(x) {
    lapply(strwrap(x, width = width, simplify = FALSE), paste, collapse="\n")}}
## colour palett
cb_pal <- c("#01665e", "#5ab4ac","#c7eae5", "#d8b365", 
            "#8c510a")
r <- c("total.area", "range.size", "effective.mesh.size", "prop.landscape", "mean.shape.index", "perimeter.area.frac.dim")

```

## read in data and models
```{r}
metrics <- readRDS("Data_metrics_for_hf_analysis.rds")
clean_tree <- read.tree("Data_sp_clean_tree.tre")

m_metric_hf <- readRDS("m_metric_hf.rds")
m_metric_clim <- readRDS("m_metric_clim.rds")
m_metric_clim_hf <- readRDS("m_metric_clim_hf.rds")
m_metric_clim_vel <- readRDS("m_metric_vel_clim.rds")
m_metric_clim_hf_vel <- readRDS("m_metric_vel_hf_clim.rds")

m_metric_null <- readRDS("m_metric_null.rds")
r2 <- readRDS("Data_r2_all_models.rds")
```

prep for exploring model results
```{r}
# names(metrics) <- c("species", "total.area", "range.size", "effective.mesh.size", "prop.landscape", "mean.shape.index", "perimeter.area.frac.dim", 
#                     "hf_mean", "vel_mean", "mat_mean", "mat_var_mean", "map_mean", "map_var_mean")
# 

mcmc_data <- metrics
mcmc_data$animal <- mcmc_data$species
## create comparative dataset
comp_data <- clean.data(mcmc_data, clean_tree, data.col = "animal")

f <- list()
f[["total.area"]]  <- total.area ~ hf_mean*map_mean*mat_mean*map_var_mean*mat_var_mean
f[["range.size"]] <- range.size ~ hf_mean*map_mean*mat_mean*map_var_mean*mat_var_mean
f[["effective.mesh.size"]] <-  effective.mesh.size ~ hf_mean*map_mean*mat_mean*map_var_mean*mat_var_mean
f[["prop.landscape"]] <- prop.landscape ~ hf_mean*map_mean*mat_mean*map_var_mean*mat_var_mean
f[["mean.shape.index"]] <- mean.shape.index ~ hf_mean*map_mean*mat_mean*map_var_mean*mat_var_mean
f[["perimeter.area.frac.dim"]] <- perimeter.area.frac.dim ~ hf_mean*map_mean*mat_mean*map_var_mean*mat_var_mean

```


```{r eval=FALSE, include=FALSE}
z <- "total.area"
# z <- "range.size"
# z <- 'effective.mesh.size'
# z <- "mean.shape.index"
# z <- "prop.landscape"
# z <- "perimeter.area.frac.dim"

# mod_mcmc <- m_metric_clim_hf[[z]][["hf"]][[1]]
# mod_mcmc_2 <- m_metric_clim_hf[[z]][["hf"]][[2]]
# 
# #bay_phylo_dia(mod_mcmc)
# 
# for(i in r){
#         print(summary(m_metric_clim_hf[[i]][["hf"]][[1]]))
# }

m_metric <- m_metric_clim_hf_vel
mod_mcmc <- m_metric[[z]][["hf"]][[1]]
mod_mcmc_2 <- m_metric[[z]][["hf"]][[2]]

#bay_phylo_dia(mod_mcmc)

for(i in r){
        print(summary(m_metric[[i]][["hf"]][[1]]))
}

```
## future WAIC/loo
```{r}

# check <- broom.mixed::tidy(m_metric_clim_hf[["total.area"]][["hf"]][[1]], effects = "fixed", conf.int = TRUE, conf.level = 0.95)
# 
# loo1 <- loo(check, save_psis = TRUE)
```

DIC table

## note that Geographic range fractality has a smaller sample size than the other models, hense the lower DIC
```{r}
l <- data.frame(cbind(r, c(1:6), c(1:6), c(1:6), c(1:6)))
l[,2] <- as.character(l[,2])
l[,3] <- as.character(l[,3])
l[,4] <- as.character(l[,4])
l[,5] <- as.character(l[,5])
for(i in r){
  l[,2][l$r == i] <- round(m_metric_clim[[i]][["hf"]][[1]][["DIC"]],2)
  l[,3][l$r == i] <- round(m_metric_hf[[i]][["hf"]][[1]][["DIC"]],2)
  l[,4][l$r == i] <- round(m_metric_clim_hf[[i]][["hf"]][[1]][["DIC"]],2)
  l[,5][l$r == i] <- round(m_metric_null[[i]][["hf"]][[1]][["DIC"]],2)
}
names(l) <- c("Metrics", "Climate", "Human_footprint", "Climate*Human_footprint", "null")
l[,2] <- as.numeric(l[,2])
l[,3] <- as.numeric(l[,3])
l[,4] <- as.numeric(l[,4])
l[,5] <- as.numeric(l[,5])
```

```{r}
l <- data.frame(cbind(r, c(1:6), c(1:6), c(1:6), c(1:6), c(1:6), c(1:6)))
l[,2] <- as.character(l[,2])
l[,3] <- as.character(l[,3])
l[,4] <- as.character(l[,4])
l[,5] <- as.character(l[,5])
l[,6] <- as.character(l[,6])
l[,7] <- as.character(l[,7])
for(i in r){
  l[,2][l$r == i] <- round(m_metric_clim[[i]][["hf"]][[1]][["DIC"]],2)
  l[,3][l$r == i] <- round(m_metric_hf[[i]][["hf"]][[1]][["DIC"]],2)
  l[,4][l$r == i] <- round(m_metric_clim_hf[[i]][["hf"]][[1]][["DIC"]],2)
  l[,5][l$r == i] <- round(m_metric_clim_vel[[i]][["hf"]][[1]][["DIC"]],2)
  l[,6][l$r == i] <- round(m_metric_clim_hf_vel[[i]][["hf"]][[1]][["DIC"]],2)
  l[,7][l$r == i] <- round(m_metric_null[[i]][["hf"]][[1]][["DIC"]],2)
}
names(l) <- c("Metrics", "Climate", "Human_footprint", "Climate*Human_footprint", "Climate*Velocity", "Climate*Vel*HF", "null")
l[,2] <- as.numeric(l[,2])
l[,3] <- as.numeric(l[,3])
l[,4] <- as.numeric(l[,4])
l[,5] <- as.numeric(l[,5])
l[,6] <- as.numeric(l[,6])
l[,7] <- as.numeric(l[,7])
```


```{r eval=FALSE, include=FALSE}
#https://haozhu233.github.io/kableExtra/awesome_table_in_html.html for colour by cell value

dic_table <- l

dic_table$Delta_DIC <- 0

for(i in 1:6){
  dic_table$Delta_DIC[i] <- dic_table[i,7] - min(dic_table[i,2:6])
  }

for(i in 2:4){
dic_table[i,6] <- paste("**", dic_table[i,6], "**", sep = "")
}
dic_table[1,5] <- paste("**", dic_table[1,5], "**", sep = "")
dic_table[5,7] <- paste("**", dic_table[5,7], "**", sep = "")
dic_table[6,7] <- paste("**", dic_table[6,7], "**", sep = "")

for(i in 1:4){
dic_table[,i] <- as.character(dic_table[,i])}

dic_table$Metrics <- c("Occupied area", "Geographic range size", "Patch size distribution", "Geographic range filling", "Patch shape complexity", "Geographic range fractality")

dic_table <- as_tibble(dic_table)
dic_table  %>%
  knitr::kable(booktabs = T) %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  pack_rows("range size", 1, 2) %>%
  pack_rows("range division", 3, 4) %>% 
  pack_rows("patch shape", 5, 6)


# dic_table 
```


## table of model results for plotting
```{r}
# sum <- list()
# for(i in r) {
# sum[[i]] <- as.data.frame(summary(m_metric_clim_hf[[i]][["hf"]][[1]][["Sol"]])[["statistics"]]);
# sum[[i]] <- setDT(sum[[i]], keep.rownames = TRUE)[]
# sum[[i]]$ci <- sum[[i]]$SD*qnorm(.95)
# sum[[i]] <- sum[[i]][2:6,] 
# sum[[i]]$rn <- c("MAT", "MAT_var", "MAP", "MAP_var", "HF")
# sum[[i]]$rn <- factor(sum[[i]]$rn,
#                       levels = c("MAP", "MAP_var", "MAT", "MAT_var", "HF"))
# }


sum <- list()
for(i in r) {
sum[[i]] <- as.data.frame(summary(m_metric_clim_hf_vel[[i]][["hf"]][[1]])$solutions);
sum[[i]] <- setDT(sum[[i]], keep.rownames = TRUE)[]
sum[[i]] <- sum[[i]][c(2:7),]
sum[[i]]$rn <- c( "Vel","HF", "MAP", "MAT", "MAP_var","MAT_var")
sum[[i]]$rn <- factor(sum[[i]]$rn,
                     levels = c( "MAP_var","MAT_var","MAP", "MAT","Vel",  "HF"))
}


## woaw....
#check <- broom.mixed::tidy(m_metric_clim_hf_vel[["total.area"]][["hf"]][[1]], effects = "fixed", conf.int = TRUE, conf.level = 0.95)

```

## plot effect sizes

```{r fig.width=15, fig.height=19}

n <- as.data.frame(cbind(c("Occupied area", "Geographic range size", "Patch size distribution", "Geographic range filling", "Patch shape complexity", "Geographic range fractality"),
           c("total.area", "range.size", "effective.mesh.size", "prop.landscape", "mean.shape.index", "perimeter.area.frac.dim")))

par(mfrow = c(6,1), mar=c(4.5,4.5,2,2), col="black", col.main = "black", col.lab = "black")
p <- list()
for(i in r){
  p[[i]] <-  ggplot(sum[[i]],  aes(rn, post.mean, colour = rn)) + 
      geom_hline(yintercept= 0, alpha = 0.8) +
  geom_line() +
  geom_pointrange(data = sum[[i]],
                    aes(rn, post.mean, colour = rn, ymin= `l-95% CI`, ymax= `u-95% CI`), 
                    position = position_dodge(0.5), size = 1.5) +
  
    theme(axis.line = element_line(colour = 'black', size = 1.5), 
          axis.title.x = element_blank(),
          axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
           plot.title = element_text(hjust = 0.5, size = 20)) +
    
  scale_x_discrete(labels = get_wraper(10)) +
  labs(colour = "Climate and HF", y = paste("Effect Size", n$V1[n$V2 ==i], sep = " ")) +
  scale_color_viridis(discrete = TRUE) +
    ylim(-0.8, 0.8) }

p[[1]] <- p[[1]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) 
p[[2]] <- p[[2]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) 
p[[3]] <- p[[3]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) 
p[[4]] <- p[[4]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank())  

resp <-ggarrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]],p[[6]], labels = c("A", "B", "C", "D", "E", "F"), nrow = 3, ncol = 2, common.legend = TRUE, legend = "bottom")
# annotate_figure(resp, top = text_grob("Effects of climate and human footprint on range metrics", #color = "black", 
#                                       face = "bold", size = 20)
resp
```

```{r}

## interactions are grey
# sum <- list()
# for(i in r) {
# sum[[i]] <- as.data.frame(summary(m_metric_clim_hf_vel[[i]][["hf"]][[1]])$solutions);
# sum[[i]] <- setDT(sum[[i]], keep.rownames = TRUE)[]
# sum[[i]] <- sum[[i]][c(2:22),]
# sum[[i]]$main <- sum[[i]]$rn
# sum[[i]]$main[grep(":", sum[[i]]$rn)] <- "interaction"
# sum[[i]]$main[sum[[i]]$pMCMC <= 0.05 & sum[[i]]$main =="interaction"] <- "sig interaction"
# sum[[i]]$main <- factor(sum[[i]]$main,
#                      levels = c("map_mean", "map_var_mean", "mat_mean", "mat_var_mean","vel_mean", "hf_mean","sig interaction", "interaction"))
# 
# }

## interactions are two coloured
sum <- list()
for(i in r) {
sum[[i]] <- as.data.frame(summary(m_metric_clim_hf_vel[[i]][["hf"]][[1]])$solutions);
sum[[i]] <- setDT(sum[[i]], keep.rownames = TRUE)[]
sum[[i]] <- sum[[i]][c(2:22),]
sum[[i]]$main <- sum[[i]]$rn
sum[[i]]$main <-  gsub("_mean", "", sum[[i]]$main, ignore.case = TRUE)
sum[[i]]$rn <-  gsub("_mean", "", sum[[i]]$rn, ignore.case = TRUE)
sum[[i]]$rn <- factor(sum[[i]]$rn,
                       levels = c("vel", "map", "mat", "map_var", "mat_var", "hf", 
                                   "hf:map","hf:mat","hf:map_var","hf:mat_var", "vel:hf",
                                  "vel:map","vel:mat","vel:map_var", "vel:mat_var",
                                  "map:mat", "map:map_var", "map:mat_var",
                                  "mat:map_var", "mat:mat_var", "map_var:mat_var"))
## hash out next 3 lines to keep non-significant terms in colour
sum[[i]]$main <- as.character(sum[[i]]$main)
sum[[i]]$main[sum[[i]]$pMCMC > 0.05] <- "ns"
sum[[i]]$main <- factor(sum[[i]]$main)
sum[[i]]$main1 <-  gsub(":.*", "", sum[[i]]$main, ignore.case = TRUE)
sum[[i]]$main2 <-  gsub(".*:", "", sum[[i]]$main, ignore.case = TRUE)
}
```


```{r fig.width=14, fig.height=18}
par(mfrow = c(6,1), mar=c(4.5,4.5,2,2), col="black", col.main = "black", col.lab = "black")

p <- list()

for(i in r){
  
p[[i]] <-  ggplot(sum[[i]],  aes(rn, post.mean, colour = main)) + 
      geom_hline(yintercept= 0, alpha = 0.8) +
  geom_line() +
  geom_pointrange(data = sum[[i]],shape = 1,
                    aes(rn, post.mean, colour = main1, ymin= `l-95% CI`, ymax= `u-95% CI`), 
                    position = position_dodge(0.5), size = 1.5) +
    geom_point(data = sum[[i]], 
                    aes(rn, post.mean, colour = main2),
                    position = position_dodge(0.5), size = 6) +        
    scale_color_manual(values = c("map" = "#440154FF",
                                  "map_var" = "#3B528BFF",
                                  "mat" = "#21908CFF",
                                  "mat_var" = "#5DC863FF",
                                  "vel" =  "orange", 
                                  "hf" = "#FDE725FF", 
                                  "ns" = "light grey")) +

    theme(axis.line = element_line(colour = 'black', size = 1.5), 
          axis.title.x = element_blank(),
          axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
           plot.title = element_text(hjust = 0.5, size = 20)) +
    
  scale_x_discrete(labels = get_wraper(10)) +
  labs(colour = "Model terms", y = paste("Effect Size", n$V1[n$V2 ==i], sep = " ")) +
    ylim(-1.2, 0.8) }

for(i in r){
  p[[i]] <- p[[i]] 
}

p[[1]] <- p[[1]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) 
p[[2]] <- p[[2]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) 
p[[3]] <- p[[3]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) 
p[[4]] <- p[[4]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank())

p[[5]] <- p[[5]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank())  
p[[6]] <- p[[6]] + theme(axis.text.x = element_blank(), axis.title.x = element_blank())  


resp <-ggarrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]],p[[6]], labels = c("A", "B", "C", "D", "E", "F"), nrow = 3, ncol = 2, common.legend = TRUE, legend = "bottom")
# annotate_figure(resp, top = text_grob("Effects of climate and human footprint on range metrics", #color = "black", 
#                                       face = "bold", size = 20)
resp

```


## pairs plots for collinearity
```{r fig.width=10, fig.height=10}
colfunc <- colorRampPalette(c("#A715AD","white", "#FFB000"))


# cols <- as.character(c(1:66))
# cols[] <- "black"
# ## get correlation matrix
# dta.r = abs(cor(rat_[, which(names(rat_) %in% c("total.area", "range.size", "effective.mesh.size",
#                                       "mean.shape.index", "prop.landscape",
#                                       "perimeter.area.frac.dim", 
#                                       "mean", "mat_mean", "mat_var_mean", 
#                                       "map_mean", "map_var_mean"))])) 
# # get colors for these correlations
# dta.col <- dmat.color(dta.r, colors = colfunc(100), byrank = NULL, breaks = 100) 
# 
# cpairs(rat_[, which(names(rat_) %in% c("total.area", "range.size", "effective.mesh.size",
#                                       "mean.shape.index", "prop.landscape",
#                                       "perimeter.area.frac.dim", 
#                                       "mean", "mat_mean", "mat_var_mean", 
#                                       "map_mean", "map_var_mean"))], 
#        panel.colors=dta.col, gap=0.5,lower.panel = NULL)



upper.panel<-function(x, y){
        r <- round(cor(x, y), digits=2)
       # points(x,y, pch=21, col = metrics$mat_var_col, cex = 0.5)
        points(x,y, pch=21, col=c("grey"), cex = 0.5)
       # points(x,y, pch=21, col=colfunc(100)[r], cex = 0.5)
        txt <- paste0("R = ", r)
       # text(0.8, 0.9, txt, cex = 6* abs(cor(x, y)))
        usr <- par("usr"); on.exit(par(usr))
        par(usr = c(0, 1, 0, 1))
        text(0.6, 0.9, txt, cex =1.8)
}

#m <- drop_na(metrics)
m <- metrics[which(names(metrics) %nin% "Geographic range fractality")]
# pairs(rat_[, which(names(rat_) %in% c("total.area", "range.size", "effective.mesh.size",
#                                       "mean.shape.index", "prop.landscape",
#                                       "perimeter.area.frac.dim", 
#                                       "mean", "mat_mean", "mat_var_mean", 
#                                       "map_mean", "map_var_mean"))],
#       lower.panel = NULL, upper.panel = upper.panel)

# pairs(m[, which(names(m) %in% c("Occupied area", "Geographic range size", "Patch size distribution", "Geographic range filling", "Patch shape complexity", "Geographic range fractality"),)],
#       lower.panel = NULL, upper.panel = upper.panel)

pairs(m[, which(names(m) %in% c("Occupied area", "Geographic range size", "Patch size distribution", "Geographic range filling", "Patch shape complexity"),)],
      lower.panel = NULL, upper.panel = upper.panel)

pairs(m[, which(names(m) %in% c("hf_mean", "vel_mean", "mat_mean", "mat_var_mean", 
                                      "map_mean", "map_var_mean"))],
      lower.panel = NULL, upper.panel = upper.panel)

                                      
```

## visualise collinearity
```{r}
# par(mfrow = c(1,3), pch = 20, cex = 1)
# 
# rbPal <- colorRampPalette(c('red','blue'))
# rbPal <- colorRampPalette(v <- viridis(400, alpha = 1, begin = 0, end = 1, direction = 1, option = "D"))
# 
# metrics$mean_col <- rbPal(820)[as.numeric(cut(metrics$hf_mean,breaks = 820))]
# metrics$mat_col <- rbPal(820)[as.numeric(cut(metrics$mat_mean,breaks = 820))]
# metrics$t.a_col <- rbPal(820)[as.numeric(cut(metrics$`Occupied area`, breaks = 820))]
# 
# plot(metrics$`Occupied area` ~ metrics$mat_mean, col = metrics$mean_col)
# plot(metrics$`Occupied area` ~ metrics$hf_mean, col = metrics$mat_col)
# plot(metrics$mat_mean ~ metrics$hf_mean, col = metrics$t.a_col)
```



## plot r2s
```{r eval=FALSE, fig.height=6, fig.width=8, include=FALSE}
r2$model <- factor(r2$model, levels = c("null","hf", "clim", "clim_hf", "vel_clim", "vel_hf_clim"))
short <- r2[r2$r == "total.area",]

 ggplot(short,  aes(model, mode_r2_mar, colour = model)) + 
  geom_pointrange(data = short,
                    aes(model, mode_r2_mar, colour = model, ymin= `lower_r2_mar`, ymax= `upper_r2_mar`), 
                    position = position_dodge(width = 0.5), size = 2) +
     geom_pointrange(data = short,
                    aes(model, mean_r2_cond, colour = model, ymin= `lower_r2_cond`, ymax= `upper_r2_cond`),
                    position = position_dodge(width = 0.5), size = 2, shape = 1) +
  
    theme(axis.line = element_line(colour = 'black', size = 1.5), 
          axis.title.x = element_blank(),
          axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
           plot.title = element_text(hjust = 0.5, size = 20)) +
  labs(colour = "Model", title = paste(i), y = "R2") +
  scale_color_viridis(discrete = TRUE) +
    ylim(-0.01, 1) 

```


```{r}
 # show_col(viridis_pal(option = "D", direction = 1)(5))
 # show_col(viridis_pal(option = "D", direction = 1)(10))
 dput(viridis_pal(option = "D", direction = 1)(5))
```

```{r}
full_pal <-c("#FA9E3BFF", "#ED7953FF","#D8576BFF", "#BD3786FF","#7301A8FF",  "#47039FFF")
```

##plot conditional only
```{r eval=FALSE, include=FALSE}
r2 <- merge(n, r2, by.x = "V2", by.y = "r")
names(r2) <- c("r_old", "r", "mean_r2_mar", "mode_r2_mar", "lower_r2_mar", "upper_r2_mar", 
              "mean_r2_cond", "mode_r2_cond", "lower_r2_cond", "upper_r2_cond", "model")
r2$model <- factor(r2$model, levels = c("null","hf", "clim", "clim_hf", "vel_clim", "vel_hf_clim"))
r2$r <- factor(r2$r, levels = c("Occupied area", "Geographic range size", "Patch size distribution", "Geographic range filling", "Patch shape complexity", "Geographic range fractality"))
r2 <- r2[r2$r %nin% c("Patch shape complexity", "Geographic range fractality"),]
 ggplot(r2,  aes(r, mode_r2_mar, colour = model)) + 
     geom_pointrange(data = r2,
                    aes(r, mean_r2_cond, colour = model, ymin= `lower_r2_cond`, ymax= `upper_r2_cond`),
                    position = position_dodge(width = 0.5), size = 2) +
  
    theme(axis.line = element_line(colour = 'black', size = 1.5), 
          axis.title.x = element_blank(),
          axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +#, plot.title = element_text(hjust = 0.5, size = 20)) +
  labs(colour = "Model", y = "R2") +
  scale_color_manual(values=paste(full_pal)) +
    ylim(-0.01, 0.75) 
```




#FDBA00 - yellow
#A302B7 - purple
#02AFA8 - turquoise

```{r fig.width=20, fig.height=10}
#r2 <- merge(n, r2, by.x = "V2", by.y = "r")
names(r2) <- c("r_old", "r", "mean_r2_mar", "mode_r2_mar", "lower_r2_mar", "upper_r2_mar", 
              "mean_r2_cond", "mode_r2_cond", "lower_r2_cond", "upper_r2_cond", "model")
r2$model <- factor(r2$model, levels = c("null","hf", "clim", "clim_hf", "vel_clim", "vel_hf_clim"))
r2$r <- factor(r2$r, levels = c("Occupied area", "Geographic range size", "Patch size distribution", "Geographic range filling", "Patch shape complexity", "Geographic range fractality"))
 ggplot(r2,  aes(r, mode_r2_mar, colour = model)) + 
  geom_pointrange(data = r2,
                    aes(r, mean_r2_mar, colour = model, ymin= `lower_r2_mar`, ymax= `upper_r2_mar`),
                    position = position_dodge(width = 0.5), size = 2) +
     geom_pointrange(data = r2,
                    aes(r, mean_r2_cond, colour = model, ymin= `lower_r2_cond`, ymax= `upper_r2_cond`),
                    position = position_dodge(width = 0.5), size = 2, shape = 1) +

   
    theme(axis.line = element_line(colour = 'black', size = 1.5), 
          axis.title.x = element_blank(),
          axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  labs(colour = "Model", y = "R2") +
  scale_color_manual(values=paste(full_pal)) +
        scale_shape_manual(values = c(1, 24), 
                         breaks = c("Conditional", "Marginal"),
                         guide = guide_legend(override.aes = list(shape = c(NA, 1), color = "black"))) +
    ylim(-0.01, 0.75) 
```

## r2 table 
```{r}
rt <- r2[,c("r", "mean_r2_mar", "mean_r2_cond", "model")]
rt <- rt[order(rt$r),] 

a <- data.frame(cbind(levels(rt$r), "mar"))
b <- data.frame(cbind(levels(rt$r), "cond"))

for(i in levels(rt$model)){
  a <- cbind(a,round(rt$mean_r2_mar[rt$model == i],3))
  b <- cbind(b,round(rt$mean_r2_cond[rt$model == i],3))}

names(a) <- c("Metrics","type", "null", "hf", "clim", "clim_hf", "vel_clim", "vel_hf_clim")
names(b) <- c("Metrics","type", "null", "hf", "clim", "clim_hf", "vel_clim", "vel_hf_clim")
rt <- rbind(a,b)

rt$Metrics <- factor(rt$Metrics, levels = c("Occupied area", "Geographic range size", "Patch size distribution", 
                          "Geographic range filling", "Patch shape complexity", "Geographic range fractality"))
rt <- rt[order(rt$Metrics),]
```



```{r}
r2_table <- as_tibble(rt)
r2_table  %>%
  knitr::kable(booktabs = T, col.names = c("Metrics","type", "null", "hf", "clim", "clim_hf", "vel_clim", "vel_hf_clim")) %>%
  add_header_above(c()) %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>% 
    collapse_rows(columns = 1:2, valign = "top") %>% 
    pack_rows("range size", 1, 4) %>%
  pack_rows("range division", 5, 8) %>%
  pack_rows("patch shape", 9, 12) 

```


```{r}
# r2_table  %>%
#   kbl(booktabs = T, col.names = c("Metrics","type", "null", "hf", "clim", "clim_hf", "vel_clim", "vel_hf_clim")) %>%
#   kable_styling(bootstrap_options = "striped", full_width = F) %>% 
#     collapse_rows(columns = 1:2, valign = "top")
```


other option for r2 plot
```{r eval=FALSE, fig.height=10, fig.width=16, include=FALSE}
r2$r <- factor(r2$r, levels = c("Occupied area", "Geographic range size", "Patch size distribution", "Geographic range filling", "Patch shape complexity", "Geographic range fractality"))
r2$model <- factor(r2$model, levels = c("null","hf", "clim", "clim_hf", "vel_clim", "vel_hf_clim"))
ggplot(r2,  aes(model, mode_r2_mar, colour = r)) + 
  geom_pointrange(data = r2,
                    aes(model, mean_r2_mar, colour = r, ymin= `lower_r2_mar`, ymax= `upper_r2_mar`), 
                    position = position_dodge(width = 0.5), size = 2) +
     geom_pointrange(data = r2,
                    aes(model, mean_r2_cond, colour = r, ymin= `lower_r2_cond`, ymax= `upper_r2_cond`),
                    position = position_dodge(width = 0.5), size = 2, shape = 1) +
  
    theme(axis.line = element_line(colour = 'black', size = 1.5), 
          axis.title.x = element_blank(),
          axis.text.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) +
  labs(colour = "Model", y = "R2") +
  scale_color_manual(values=paste(full_pal)) +
  #scale_color_viridis(discrete = TRUE) +
    ylim(-0.01, 0.75) 
```
















